### 2017-02-04 hpc_101/sections/09_related_services.tex

\section{Related Services}
\begin{frame}
\frametitle{Services}
\begin{itemize}
\item MATLAB Distributed Computing Server{\trademark}
  \begin{itemize}
    \item Requires user to already have a license for MATLAB
    \item User must also have access to at least the Parallel Computing Toolbox{\trademark}
    \item Provides users access to up to 64 workers
    \item Requires some additional configuration
    \item Email us to request access to this additional service
  \end{itemize}
\end{itemize}
\end{frame}