### 2018-04-03 hpc_101/sections/03_sustainability.tex

\section{Sustainability}
\subsection{Community Resources}
\begin{frame}
	\frametitle{Community Resources}
	\begin{itemize}
		\item The initial investment in the cluster provides resources for free to all faculty, staff, and students affiliated with the University of {\hawaii}
		\item All users can run on the publicly accessible partitions:\\ \emph{community.q, exclusive.q, lm.q, gpu.q, sb.q, kill.q, kill.gpu.q, htc.q}
		\item For some users, the publicly available resources may not be enough~\ldots
	\end{itemize}
\end{frame}

\subsection{Ownership Models}

\subsubsection{Condo Model}
\begin{frame}
	\frametitle{Condo Model}
	\begin{itemize}
		\item The Condo model allows users to buy nodes (\emph{condos}) to incorporate into the cluster	
		\item Node owners are provided with priority access to their purchased hardware
%		\item All nodes have a \emph{5 year warranty}
%		\begin{itemize}
%			\item Once a nodes warranty has expired, it will be removed from the cluster
%		\end{itemize}
%		\item Condo owners are given early access to purchased nodes
%		  \begin{itemize}
%                    \item Access is granted as soon as the funds land in our accounts
%		    \item The 5 year warranty will not begin until newly ordered nodes are installed
%		\end{itemize}
		\item Condo owners are also given the option to purchase~\\1TB of {\lustre} storage per node purchased
                \item {\textcolor{red}{\textbf{ Condo purchases currently suspended} }}
	\end{itemize}
        \btVFill
        \begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
\end{frame}

\subsubsection{Node Leasing}
\begin{frame}
	\frametitle{Leasing}
	\begin{itemize}
        \item In some cases, users will not require owning a node for multiple years, but still need priority access
	\item Leases have a 1 year term
        \item Leases are $\$1,320$ per year, which is \( \frac{1}{5} \) the cost of a condo purchase
        \item Node leasees are provided priority access to their leased hardware
        \item Leases are not considered an equipment purchase
        \item Total availabity limited to $\approx100$ community nodes
	\end{itemize}
        \btVFill
        \begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
\end{frame}

\subsubsection{Service Units}
\begin{frame}	
	\frametitle{Service Units}
	\begin{itemize}
	 \item In other cases, for users that require a large burst of dedicated nodes, service units (\emph{SU}) is an option
	 \item SUs come in two flavors:
		 \begin{itemize}
			 \item[--] Standard node units -- (\$0.50 per SU) 
				\begin{itemize}
					\item 20 core hours, with access to 128GB of ram on a standard~node
				\end{itemize}
			 \item[--] Large memory node units -- (\$2.00 per SU)
				\begin{itemize}
					\item 40 core hours, with access to 1TB of ram on a large~memory~node
				\end{itemize}
		 \end{itemize}
	 \item A minimum order totaling \$500 is required
	\end{itemize}
        \btVFill
	\begin{center}{\footnotesize \emph{All prices are subject to change}}\end{center}
\end{frame}

%% \subsubsection{Price Card}
%% \begin{frame}
%% 	\frametitle{Price Card}
%% 	\begin{center}
%% 	{\large \textbf{Condo Price Card}} \\
%% 	\resizebox{\textwidth}{!}{%
%% 		\begin{tabular}{l r ||l r}
%% 			\toprule
%% 			\textbf{Product} & \textbf{Cost} & \textbf{Product} & \textbf{Cost} \\
%% 			\midrule
%% 			\midrule
%% 			Standard & \$6,600.00  & GPU Node& Contact for pricing\\
%% 			\hline
%% 			Large Memory & \$20,000.00  & 1 TB {\lustre} Storage for 5 years & \$$600.00$ \\
%% 			\bottomrule
%% 		\end{tabular}
%% 		}
%% 	\end{center}
%% 	\smallskip
%% 	\begin{center}
%% 		{\large \textbf{Service Unit Price Card}}\\
%% 		\begin{tabular}{l || c || r}
%% 			\toprule
%% 			\textbf{Product} & \textbf{Cost} & \textbf{Minimum Order}\\
%% 			\midrule
%% 			\midrule
%% 			Standard node &  &  1,000~SU (\$500.00)\\
%% 			Large memory node & \$2.00 per SU & 250~SU (\$500.00)\\
%% 			\bottomrule
%% 		\end{tabular}
%% 		\btVFill
%% 	{\footnotesize \emph{All prices are subject to change}}
%% 	\end{center}
%% \end{frame}