### 2020-11-25 OOD/sections/02_walking.tex
Update to slides to point to caltech j.notebook instructions.  Merged in changes made for a class specific version to include where to contact us for support.

\part{Connecting \& Tour}
\begin{frame}
			 \partpage
\end{frame}

\section[Connecting]{Prerequisite \& Connecting}
\begin{frame}
\frametitle{Prerequisite}
		\begin{block}{Prerequisite}
			\begin{itemize}
				\item An active Mana (UH-HPC) account
				\begin{itemize}
				\item Accounts provisioned for a class are active until the end of the semester
				\end{itemize}
				\item Registered for \href{http://www.hawaii.edu/its/uhlogin/}{MFA/DUO}
				\item A web browser 
			\end{itemize}
			\end{block}
			\begin{exampleblock}{Supported Browsers}
			\begin{itemize}
				\item Chrome\footnote[1,frame]{\tiny Chrome is the recommended browser}\footnote[2,frame]{\tiny \url{https://osc.github.io/ood-documentation/latest/requirements.html\#browser-requirements}}, Firefox, Edge are all supported
		\end{itemize}
		\end{exampleblock}
			\begin{alertblock}{Unsupported Browsers}
			\begin{itemize}
				\item Internet Explorer 11
		\end{itemize}
				\end{alertblock}
\end{frame}

\begin{frame}
\frametitle{Connecting}
			\begin{itemize}
				\item Open up a browser window in incognito mode
					\begin{itemize}
					\item Optional but recommended for security upon logout
					\end{itemize}						
				\item Direct your browser to \url{https://uhhpc.its.hawaii.edu}
				\item Should be prompted for your UH credentials by the UH gold screen
				\item If successful, should be directed to a page with the Mana logo
			\end{itemize}
	\begin{center}
  \includegraphics[scale=0.15]{05.png}
	\end{center}			
\end{frame}

\section[WT]{Tour}
\begin{frame}
\Huge{\centerline{Tour}}
\end{frame}

\section{User Support}
\begin{frame}
  \frametitle{User Support}
  \begin{block}{Contact Information}
   For help, please contact us at: UH-HPC-Help@lists.hawaii.edu
  \end{block}
  \begin{block}{Online documents \& FAQ}
		\begin{itemize}
		\item \href{http://datascience.hawaii.edu/hpc/}{Information \& policies}
		\item \href{http://go.hawaii.edu/jdG}{FAQ}
		\item \href{http://go.hawaii.edu/3A8}{Video Tutorials}
		\end{itemize}
  \end{block}
\end{frame}