### 2023-05-03 hpc_101/sections/09_appendix.tex
monitoring examples added

\part{Appendix}

\section[Terminology]{Terminology}

\begin{frame}
	\frametitle{Terminology}
	\begin{itemize}
	\item \textbf{Node} -- Another name for a server or computer
        \item \textbf{Login node} -- A specialized node that users connect to in order to submit work to a computer cluster
        \item \textbf{Computer cluster} -- A set of loosely or tightly connected nodes that work together so that, in many respects, they can be viewed as a single system\footnote{\label{wiki_ccluster}\tiny\url{https://en.wikipedia.org/wiki/Computer_cluster}}
        \item \textbf{Data transfer node (DTN)} -- Specialized nodes that minimize the impedance on the network to access the full capability of the network
	\item \textbf{Science DMZ (SciDMZ)} --  A portion of the network configured with equipment and security policies in order to optimize for high-performance scientific applications rather than for general-purpose business systems or “enterprise” computing 
	\item \textbf{Multi-factor Authentication (MFA)} -- An authentication method in which a computer user is granted access only after successfully presenting two or more pieces of evidence or factors to an authentication mechanism, e.g., DUO

	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Terminology}
	\begin{itemize}
	\item \textbf{Symbolic Link (symlink)} -- A file that contains a reference to another file or directory
	\item \textbf{Command-line interface/interpreter (CLI)} -- A text-based user interface used to view and manage computer files 
	\item \textbf{Message Passing Interface (MPI)} -- A standard that is used by programs to pass messages between nodes
	\item \textbf{High Performance Compute (HPC)} -- A computing paradigm in which applications are typically a tightly coupled parallel job that benefit from a low-latency interconnect
	\item \textbf{High Throughput Compute (HTC)} -- A computing paradigm that focuses on the efficient execution of a large number of loosely-coupled tasks
        \item \textbf{Modified time} -- The last time the file was modified (content has been modified)\footnote{\label{Types of timestamps}\tiny\url{https://unix.stackexchange.com/a/2465}}
        \item \textbf{Shell script (script)} -- A computer program designed to be run by a CLI\footnote{\label{Shell script}\tiny\url{https://en.wikipedia.org/wiki/Shell_script}}
%        \item Pleasantly Parallel -- Processes are independent and no communication is necessary 
	\end{itemize}

\end{frame}
\section{Monitoring Jobs and Partitions}{Monitoring Jobs and Partitions}
\begin{frame}[fragile]
\frametitle{Monitoring Running Jobs}
\begin{block}{\textbf{squeue}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ squeue --me
             JOBID PARTITION     NAME     USER ST       TIME  NODES NODELIST(REASON)
          43486015   sandbox     bash victorgc  R    2:05:37      1 node-0093
\end{semiverbatim}
\end{block}
The --me flag can be removed and \textbf{squeue} will report all running jobs in the system.
\begin{block}{\textbf{squeue}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ squeue | head
             JOBID PARTITION     NAME     USER ST       TIME  NODES NODELIST(REASON)
          43078226   bioinfo    sgSQL     user  R 23-02:12:35      1 node-0327
          43078227   bioinfo cromwell     user  R 23-02:12:09      1 node-0327
          43485890   bioinfo       QC     user  R    4:45:59      1 node-0327
          43399271   bioinfo      TQC     user  R 3-20:49:10      1 node-0328
          43485868 exclusive    MoSe2     user  R    5:20:14      1 node-0296
          43485867 exclusive    MoSe2     user  R    5:30:46      1 node-0295
          43485866 exclusive    MoSe2     user  R    5:35:43      1 node-0294
          43485815 exclusive    MoSe2     user  R    5:57:03      1 node-0281
\end{semiverbatim}
\end{block}
\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/squeue.html}{https://slurm.schedmd.com/squeue.html}}

\end{frame}

\begin{frame}[fragile]
\frametitle{Completed Jobs Reports}
\begin{block}{\textbf{sacct}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ sacct -u victorgc -X | head
JobID           JobName  Partition    Account  AllocCPUS      State ExitCode
------------ ---------- ---------- ---------- ---------- ---------- --------
43486049_1     spectrum     shared         uh          1  COMPLETED      0:0
43486049_12    spectrum     shared         uh          1  COMPLETED      0:0
\end{semiverbatim}
\end{block}
\begin{block}{\textbf{seff}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ seff 43486049_12
Job ID: 43486061
Array Job ID: 43486049_12
Cluster: uh_hpc
User/Group: victorgc/victorgc
State: COMPLETED (exit code 0)
Cores: 1
CPU Utilized: 00:00:33
CPU Efficiency: 97.06\% of 00:00:34 core-walltime
Job Wall-clock time: 00:00:34
Memory Utilized: 75.91 MB
Memory Efficiency: 1.27\% of 5.86 GB
\end{semiverbatim}
\end{block}
The output of \textbf{sacct} is highly customizable.
\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/sacct.html}{https://slurm.schedmd.com/sacct.html}}

\end{frame}

\begin{frame}[fragile]
\frametitle{Partition Information}
\begin{block}{\textbf{sinfo}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ sinfo | head
PARTITION       AVAIL  TIMELIMIT  NODES  STATE NODELIST
gpu                up 3-00:00:00      9    mix gpu-[0013,0017-0024]
gpu                up 3-00:00:00      1  alloc gpu-0016
sandbox            up    4:00:00      1    mix node-0093
sandbox            up    4:00:00      3  alloc node-[0094-0096]
shared*            up 3-00:00:00      4  down* node-[0102,0120,0122,0178]
shared*            up 3-00:00:00      8  drain node-[0111-0112,0126,0130,0150,0175,0344,0350]
\end{semiverbatim}
\end{block}
Partition information can be printed in a per node basis.
\begin{block}{\textbf{sinfo}}
\begin{semiverbatim}\tiny \texttt
[victorgc@node-0093 HDM]\$ sinfo -Nel | head
Wed May 03 01:20:33 2023
NODELIST   NODES       PARTITION       STATE CPUS    S:C:T MEMORY TMP_DISK WEIGHT AVAIL_FE REASON
gpu-0001       1     gpu-sandbox       mixed 20     2:10:1 128500        0  10000 x86,inte none
gpu-0001       1  kill-exclusive       mixed 20     2:10:1 128500        0  10000 x86,inte none
gpu-0001       1     kill-shared       mixed 20     2:10:1 128500        0  10000 x86,inte none
gpu-0002       1  kill-exclusive        idle 20     2:10:1 128500        0  10000 x86,inte none
gpu-0002       1     kill-shared        idle 20     2:10:1 128500        0  10000 x86,inte none
\end{semiverbatim}
\end{block}
\numlessfootnotetxt{\tiny \href{https://slurm.schedmd.com/sinfo.html}{https://slurm.schedmd.com/sinfo.htmls}}
\end{frame}